---
title: "Process Files"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

To keep these data anonymous, I' processing the source files to remove GPS coordinates. I could simply set coordinates to 0, but I like to work with metric distances, if nothing else than to make it easier to think of yield in terms compatible with machine dimensions.

I do, however, retain the historical family names. These are typically the names associated with a family who owned or farmed a particular cropland in the past. The files were typically exported from `myjohndeere.deere.com` with longer file names. In the past, the data included 7 `channels` per sample, but more recent exports aggregate time samples. I've kept some of the older file formats in `Channels` folders on my machine. The files in these projects should not have channel data.

I will also rename some fields to make the data fields more self-documenting.

```{r}
file.names <- c("")
file.path <- c("~/Work/git/statistics/test_cases/claussen/")

source('~/Work/git/statistics/R/add.metric.R')
source('~/Work/git/statistics/R/plot.gps.grids.R')
source('~/Work/git/statistics/R/aggregate.seed.fn.R')
source('~/Work/git/statistics/R/aggregate.harvest.fn.R')
```

## Shapefiles

```{r,eval=FALSE}
field.name <- 'Dave Claussen-Kenny-cash-AndersonsStevens-Soybeans'
field.name <- 'Dave Claussen-Kenny-cash-AndersonsStevensW-Corn.shp'
file.name <-paste('~/Work/git/statistics/test_cases/claussen/Harvest/2013',field.name,sep='/')

#library(maptools)
#getinfo.shape(file.name)
#fieldA.2016dat <- readShapeSpatial(file.name)
library(sf)
fieldA.2016dat <- st_read(file.name)
summary(fieldA.2016dat)
#plot(fieldA.2016dat["VRYIELDVOL"])
head(fieldA.2016dat["VRYIELDVOL"])
m.sf = st_as_sf(fieldA.2016dat["VRYIELDVOL"])
plot(m.sf)
head(as(fieldA.2016dat["VRYIELDVOL"],'Spatial'))
#library(shapefiles)
#shapefile <- read.shapefile(file.name)
```




```{r}
library(ggplot2)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#F0E442","#CC79A7","#000000","#734f80", "#2b5a74", "#004f39", "#787221", "#003959", "#6aaf00", "#663cd3")

```

# Data

These data have multiple channels per harvest pass (the combine header is divided into sections), so we need to combine yield observations. This section reads and processes the data. 

Two utility functions to automate the first step of aggregation:
Note that these functions assume data are exported (from QGIS) with `X` and `Y` coordinates.

# Home

```{r}
dir <- '~/Work/git/statistics/test_cases/claussen/'
overview.path <- paste(dir,'Overview.csv',sep='')
Overview <- read.csv(overview.path)

#Use this for treated B
#Overview = Overview[Overview$Group==1,]

activities = c('Harvest','Seeding','Application')

seeding.columns <- c('Longitude','Latitude','AppliedRate','ControlRate','Variety',"DISTANCE","Heading")
harvest.columns <- c('Longitude','Latitude','Yield',"SWATHWIDTH",'Swaths',"DISTANCE","Heading","WetMass","Moisture")

groups <- vector(mode='list',length=length(unique(Overview$Group)))
tables <- vector(mode='list',length=length(unique(Overview$Group)))
names(tables) <- unique(Overview$Group)
for(g in length(unique(Overview$Group))) {
  tables[[g]] <- vector(mode='list')
}

for(i in 1:dim(Overview)[1]) {
  Crop <- Overview$Crop[i]
  if(Crop!='') {
    
    for(activity in activities) {
      path <- paste(dir,Overview$Field[i],'/',Overview$Field[i],sep='')
      path <- paste(path,Overview$Year[i],Overview$Crop[i],activity,sep=' ')
      path <- paste(path,'csv',sep='.')
      
      
      if(file.exists(path)) {
        current.dat <- read.csv(path,header=TRUE,comment.char = "#")
        if(activity=='Seeding') {
          tmp <- NULL
          try(tmp <- aggregate.seed.fn(current.dat))
          if(is.null(tmp)) {
            print(path)
          } else {
            current.dat <- tmp
            current.dat$VRYIELDVOL <- current.dat$AppliedRate
          }
        } 
        if(!is.null(current.dat)) {
         # current.dat <- add.metric(current.dat)
          current.dat$Swaths <- 7
          current.dat$SWATHWIDTH <- 5
          current.dat$Yield <- current.dat$VRYIELDVOL
          attr(current.dat,'activity') <- activity
          attr(current.dat,'crop') <- Overview$Crop[i]
          attr(current.dat,'year') <- Overview$Year[i]
          attr(current.dat,'code') <- Overview$Code[i]
          tables[[Overview$Group[i]]] <- append(tables[[Overview$Group[i]]], list(current.dat)) 
        }
      }
    }
  }
}
```

```{r}
out.dir <- './data'
b.origin = c()
for(g in 1:length(tables)) {
  current.fields <- tables[[g]]
  
  min.X <- Inf
  min.Y <- Inf
  code=''
  if(length(current.fields)>0) {
    for(f in 1:length(current.fields)) {
      if(!is.null(current.fields[[f]])) {
        current.dat <- current.fields[[f]]
        current.X <- min(current.dat$X)
        current.Y <- min(current.dat$Y)

        crop <- attr(current.dat,'crop')
        year <- attr(current.dat,'year')
        code <- attr(current.dat,'code')
        mask <- Overview$Code==code & Overview$Year==year
        
        if(Overview$X[mask] > current.X) {Overview$X[mask] <- current.X}
        if(Overview$Y[mask] > current.Y) {Overview$Y[mask] <- current.Y}
        if(current.X<min.X) {min.X <- current.X}
        if(current.Y<min.Y) {min.Y <- current.Y}
      }
    }
  }

  origin <- c(min.X,min.Y)
  
  if(length(current.fields)>0) {
  for(f in 1:length(current.fields)) {
    current.dat <- current.fields[[f]]
    if(!is.null(current.fields[[f]])) {
      current.dat <- add.metric(current.dat,origin)
      #plot.gps.grids(current.dat,plots=0,residuals=TRUE)
    
      out.activity <- attr(current.dat,'activity')
      crop <- attr(current.dat,'crop')
      year <- attr(current.dat,'year')
      code <- attr(current.dat,'code')

      out.path <-  paste(out.dir,code,sep='/')
      out.path <- paste(out.path,year,crop,out.activity,sep=' ')
      out.path <- paste(out.path,'csv',sep='.')
      print(out.path)
    
      if(out.activity=='Seeding') {
          try(write.csv(current.dat[,seeding.columns],file=out.path))
      } else {
          write.csv(current.dat[,harvest.columns],file=out.path)
      }
    }
  }
  }
}

#write.csv(Overview,file=overview.path)
```

```{r}
#soybean.2019.dat <- read.csv('~/Work/git/statistics/test_cases/claussen/AndersonsStevens/AndersonsStevens 2019 Soybeans Harvest.csv')
#B <- Overview[Overview$Group==1,]
#origin <- c(min(B$X),min(B$Y))
```

```{r,eval=FALSE}
treated.dat <- read.csv('~/Work/git/ASA_CSSA_SSSA/2019/AndersonsStevens-Soybeans.csv',header=TRUE,comment.char = "#")
#c(min(treated.dat$X),min(treated.dat$Y))
treated.dat <- add.metric(treated.dat,origin)
out.path <-  paste(out.dir,'B',sep='/')
out.path <- paste(out.path,2019,'Soybeans','Treated',sep=' ')
out.path <- paste(out.path,'csv',sep='.')
out.path
#names(current.dat)[1] <- 'Longitude'
#names(current.dat)[2] <- 'Latitude'
names(treated.dat)[7] <- 'Yield'
write.csv(treated.dat[,c('Longitude','Latitude','Time','Yield','Block','Sample','Balanced','Rectangle','Square')],file=out.path)

ggplot(treated.dat, aes(X,Y)) + 
geom_point(aes(colour = Block),size=1)
```

```{r}
ggplot(treated.dat, aes(Longitude,Latitude)) + 
geom_point(aes(colour = Block),size=1) 
```


For some reason, I have a few csv files with different column headers. I can't remember, perhaps I exported from SMS?






