---
title: "Overview"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#theme_set(theme_dag())
#since we have k-fold cross validation

set.seed(1000)
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggformula)
library(GGally)
library(ggdag)
library(mgcv)
library(bnlearn)
library(splines)

  grey <- "#999999"
  orange <- "#E69F00"
  skyblue <- "#56B4E9"
  bluishgreen <- "#009E73"
  yellow <- "#F0E442"
  blue <- "#0072B2"
  vermillion <- "#D55E00"
  
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#F0E442","#CC79A7","#000000","#734f80", "#2b5a74", "#004f39", "#787221", "#003959", "#6aaf00", "#663cd3")

remove.harvest.outliers.fn <- function(tbl,rng=3) {
  tbl <- tbl[!is.na(tbl$Yield),]
  yield.mean <- mean(tbl$Yield)
  yield.sd <- sd(tbl$Yield)
  tbl <- tbl[tbl$Yield<=yield.mean+rng*yield.sd,]
  tbl<- tbl[tbl$Yield>=yield.mean+rng-yield.sd,]
  return(tbl)
}
remove.seed.outliers.fn <- function(tbl) {
  tbl <- tbl[tbl$AppliedRate<=32000,]
  tbl<- tbl[tbl$AppliedRate>=20000,]
  return(tbl)
}
```

# Background

I assigned a project to a group of data science students. We have 5 data sets, each representing a single corn field. The data contain two columns of interest. `Yield` is measured in bushels per acre, while `ControlRate` is measured in seeds per acre. Each row in the file represents a geo-spatially tagged point located by `Easting` and `Northing`, measured in meters from the southwest corner of the field.

```{r,fig.width=8,fig.height=4.5}
aggregate.rates <- function(dat) {
  dat$Rate <- 1000*round(dat$ControlRate/1000)
  return(aggregate(Yield ~ Rate,data=dat,mean,na.rm=TRUE))
}
Rates.dat <- aggregate.rates(read.csv(file='./data/fieldA.csv'))
Rates.dat$Field <- 'D'
tmp <- aggregate.rates(read.csv(file='./data/fieldB.csv'))
tmp$Field <- 'A'
Rates.dat <- rbind(Rates.dat,tmp)
tmp <- aggregate.rates(read.csv(file='./data/fieldC.csv'))
tmp$Field <- 'C'
Rates.dat <- rbind(Rates.dat,tmp)
tmp <- aggregate.rates(read.csv(file='./data/fieldD.csv'))
tmp$Field <- 'B'
Rates.dat <- rbind(Rates.dat,tmp)
ggplot(Rates.dat, aes(Rate,Yield,color=Field)) + scale_colour_manual(values=cbPalette) +
geom_point(size=3) + geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE)
par(mfrow=c(2,2))
plot(Northing ~ Easting, data=read.csv(file='./data/fieldA.csv'),main="A",pch='.')
plot(Northing ~ Easting, data=read.csv(file='./data/fieldB.csv'),main="B",pch='.')
plot(Northing ~ Easting, data=read.csv(file='./data/fieldC.csv'),main="C",pch='.')
plot(Northing ~ Easting, data=read.csv(file='./data/fieldD.csv'),main="C",pch='.')
```

```{r,fig.width=8,fig.height=4}
aggregate.rates <- function(dat) {
  dat$Rate <- 500*round(dat$ControlRate/500)
  return(aggregate(Yield ~ Rate,data=dat,mean,na.rm=TRUE))
}
Rates.dat <- aggregate.rates(read.csv(file='./data/fieldA.csv'))
Rates.dat$Field <- 'A'
tmp <- aggregate.rates(read.csv(file='./data/fieldB.csv'))
tmp$Field <- 'B'
Rates.dat <- rbind(Rates.dat,tmp)
tmp <- aggregate.rates(read.csv(file='./data/fieldC.csv'))
tmp$Field <- 'C'
Rates.dat <- rbind(Rates.dat,tmp)
tmp <- aggregate.rates(read.csv(file='./data/fieldD.csv'))
tmp$Field <- 'D'
Rates.dat <- rbind(Rates.dat,tmp)
ggplot(Rates.dat, aes(Rate,Yield,color=Field)) + scale_colour_manual(values=cbPalette) +
geom_point(size=3) + geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE)
```


```{r,fig.width=8,fig.height=4.5}
fields <- vector(mode='list',length=5)
Fields <- read.csv(file='./data/fieldA.csv')
Fields$Field <- 'A'
tmp <- read.csv(file='./data/fieldB.csv')
tmp$Field <- 'B'
Fields <- rbind(Fields,tmp)
tmp <- read.csv(file='./data/fieldC.csv')
tmp$Field <- 'C'
Fields <- rbind(Fields,tmp)
tmp <- read.csv(file='./data/fieldD.csv')
tmp$Field <- 'D'
Fields <- rbind(Fields,tmp)
```

```{r,fig.width=8,fig.height=4.5}
Fields$Field <- factor(Fields$Field)
Fields$SeedingRate <- factor(ceiling(Fields$ControlRate/500)*500)
ggplot(Fields, aes(SeedingRate, Yield,color=Field)) +
  geom_boxplot(outlier.alpha = 0.0) +  scale_colour_manual(values=cbPalette) +facet_wrap(~ Field)
```

```{r,fig.width=8,fig.height=4.5}
ggplot(Fields, aes(ControlRate, Yield,color=Field)) +
  geom_boxplot(aes(group = cut_width(ControlRate, 500)), outlier.alpha = 0.1) +
geom_jitter(width = 100,alpha=0.1,size=0.3) + facet_wrap(~Field) + scale_colour_manual(values=cbPalette)
```


# Introduction

Suppose we have a yield map and a seeding map for a single field. The seeding map contains seeding rate estimates for discrete points in a crop land - say, a corn field, measured in seeds per acre. The yield map contains yield estimates, also associated with discrete points in a crop land, measured in bushels per acre. We wish to determine if there is a relationship between seeding rate and yield. 

More specifically, suppose this cropland has been historically managed using uniform rate methods. Seeds were planted at a constant rate (say, 28000 seeds/acre) over the entire cropland. However, for on growing season, the farmer decided to try variable-rate seeding, and changed seeding rate over different regions of the field. Can we determine if there was a positive response to seeding rate? Alternatively, can we identify an optimal seeding rate for this cropland?


# Data 

We start with yield and seeding maps. Both include GPS coordinates with associated with either yield estimates, in bushels per acre in the case of yield maps and programmed seeding rate (`ControlRate`) and actual seeding rate (`AppliedRate`) in the case of seeding maps. The GPS coordinate where recorded as the position of the seeder or harvester at one second intervals during seeding or harvest. Since these are two distinct operations, yield sample points rarely coincide with seeding sample points. To measure correlations between seeding and yield require imputation to combine the data.

In this document, we will consider two methods. In the simpler case, we impose a grid over the field and average over samples in grid cells (see `OptimalGridSize`). This has the advantage of being computationally simpler, but will be less precise. We also use inverse distance weighting (IDW) to map yield estimates to seed sample points. This may be more informative but is also noisier.



## Load Data

```{r,eval=TRUE}
harvest.2013.dat <- read.csv(file='./data/A 2013 Soybeans Harvest.csv')
harvest.2015.dat <- read.csv(file='./data/A 2015 Wheat Harvest.csv')
harvest.2018.dat <- read.csv(file='./data/A 2018 Corn Harvest.csv')
harvest.2016.dat <- read.csv(file='./data/A 2016 Corn Harvest.csv')
harvest.2017.dat <- read.csv(file='./data/A 2017 Soybeans Harvest.csv')
seed.2018.dat <-read.csv(file='./data/A 2018 Corn Seeding.csv')
harvest.2019.dat <- read.csv(file='./data/A 2019 Soybeans Harvest.csv')
seed.2020.dat <-read.csv(file='./data/A 2020 Corn Seeding.csv')
harvest.2020.dat <-read.csv(file='./data/A 2020 Corn Harvest.csv')
```

Screen for outliers.

```{r}
seed.2018.dat <- remove.seed.outliers.fn(seed.2018.dat)
seed.2020.dat <- remove.seed.outliers.fn(seed.2020.dat)
harvest.2020.dat <- remove.harvest.outliers.fn(harvest.2020.dat)
harvest.2019.dat <- remove.harvest.outliers.fn(harvest.2019.dat)
harvest.2018.dat <- remove.harvest.outliers.fn(harvest.2018.dat)
harvest.2017.dat <- remove.harvest.outliers.fn(harvest.2017.dat)
harvest.2013.dat <- remove.harvest.outliers.fn(harvest.2013.dat)
harvest.2015.dat <- remove.harvest.outliers.fn(harvest.2015.dat)
harvest.2016.dat <- remove.harvest.outliers.fn(harvest.2016.dat)
```



```{r,eval=FALSE}
harvest.2018.dat <- read.csv(file='./data/E 2018 Corn Harvest.csv')
harvest.2016.dat <- read.csv(file='./data/E 2016 Corn Harvest.csv')
harvest.2017.dat <- read.csv(file='./data/E 2017 Soybean Harvest.csv')
seed.2018.dat <-read.csv(file='./data/E 2018 Corn Seeding.csv')
```

It is easier to visualize using ranks.

```{r}
harvest.2018.dat$YieldRank <- rank(harvest.2018.dat$Yield)
harvest.2018.dat$YieldRank <- harvest.2018.dat$YieldRank/max(harvest.2018.dat$YieldRank)
seed.2018.dat$RateRank <- rank(seed.2018.dat$ControlRate)
seed.2018.dat$RateRank <- seed.2018.dat$RateRank/max(seed.2018.dat$RateRank)

seed.2020.dat$RateRank <- rank(seed.2020.dat$ControlRate)
seed.2020.dat$RateRank <- seed.2020.dat$RateRank/max(seed.2020.dat$RateRank)
```

Plot the seeding and yield maps for 2018

```{r,fig.width=8,fig.height=4.5}
Maps <- data.frame(Longitude=c(harvest.2018.dat$Longitude,seed.2018.dat$Longitude),
                   Latitude=c(harvest.2018.dat$Latitude,seed.2018.dat$Latitude),
                   Value=c(harvest.2018.dat$YieldRank,seed.2018.dat$RateRank),
                   Map=c(rep('Yield',length(harvest.2018.dat$YieldRank)),
                         rep('Seeding',length(seed.2018.dat$RateRank))))
ggplot(Maps, aes(Longitude,Latitude)) + 
geom_point(aes(colour = Value),size=0.5) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Seeding and Yield (2018)") + facet_wrap(~ Map)
```

How was yield rate determined? In this case, *ad-hoc* Rate was increased in areas where yield was estimated (from past experience) to be higher in the past; rate decreased in historically lower yield areas. So, it seems reasonable to compare with 2017 yield.

```{r}
harvest.2017.dat$YieldRank <- rank(harvest.2017.dat$Yield)
harvest.2017.dat$YieldRank <- harvest.2017.dat$YieldRank/max(harvest.2017.dat$YieldRank)
```

```{r,fig.width=8,fig.height=4.5}
Maps <- data.frame(Longitude=c(harvest.2017.dat$Longitude,seed.2018.dat$Longitude),
                   Latitude=c(harvest.2017.dat$Latitude,seed.2018.dat$Latitude),
                   Value=c(harvest.2017.dat$YieldRank,seed.2018.dat$RateRank),
                   Map=c(rep('Yield',length(harvest.2017.dat$YieldRank)),
                         rep('Seeding',length(seed.2018.dat$RateRank))))
ggplot(Maps, aes(Longitude,Latitude)) + 
geom_point(aes(colour = Value),size=1) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Seeding (2018) and Yield (2017)") + facet_wrap(~ Map)
```


```{r,fig.width=8,fig.height=4.5}
Maps <- data.frame(Longitude=c(harvest.2017.dat$Longitude,harvest.2018.dat$Longitude),
                   Latitude=c(harvest.2017.dat$Latitude,harvest.2018.dat$Latitude),
                   Value=c(harvest.2017.dat$YieldRank,harvest.2018.dat$YieldRank),
                   Map=c(rep('Yield 2017',length(harvest.2017.dat$YieldRank)),
                         rep('Yield 2018',length(harvest.2018.dat$YieldRank))))
ggplot(Maps, aes(Longitude,Latitude)) + 
geom_point(aes(colour = Value),size=0.5) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Yield 2017 and 2018") + facet_wrap(~ Map)
```

```{r,fig.width=6,fig.height=6}
Maps <- data.frame(Longitude=c(harvest.2017.dat$Longitude,harvest.2018.dat$Longitude,seed.2018.dat$Longitude),
                   Latitude=c(harvest.2017.dat$Latitude,harvest.2018.dat$Latitude,seed.2018.dat$Latitude),
                   Value=c(harvest.2017.dat$YieldRank,harvest.2018.dat$YieldRank,seed.2018.dat$RateRank),
                   Map=c(rep('Yield 2017',length(harvest.2017.dat$YieldRank)),
                         rep('Yield 2018',length(harvest.2018.dat$YieldRank)),
                         rep('Seeding 2018',length(seed.2018.dat$RateRank))))
ggplot(Maps, aes(Longitude,Latitude)) + 
geom_point(aes(colour = Value),size=0.5) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Correlation Maps") + facet_wrap(~ Map,nrow=2)
```

# Spatial Analysis

Since there is a spatial component, we'll start with a simplified spatial analysis. We'll divide the field into grid to map seeding data to yield data. 


```{r,echo=FALSE}
grid.field <- function(harvest.dat,response='Yield',grid.width=c(50,50),FUN=mean) {
  harvest.dat$Row <- ceiling(harvest.dat$Latitude/grid.width[1])
  harvest.dat$Column <- ceiling(harvest.dat$Longitude/grid.width[2])
  return(harvest.dat)
}
aggregate.field <- function(harvest.dat,response='Yield',grid.width=c(50,50),FUN=mean) {
  harvest.dat$Row <- ceiling(harvest.dat$Latitude/grid.width[1])
  harvest.dat$Column <- ceiling(harvest.dat$Longitude/grid.width[2])
  fmla <- as.formula(paste(response,'~ Row + Column'))
  tmp <- aggregate(fmla,data=harvest.dat,FUN=FUN,na.rm=TRUE)
  count <- aggregate(fmla,data=harvest.dat,FUN=length)
  row.names(tmp) <- paste(tmp$Row,tmp$Column,sep=":")
  tmp$Samples <- count[,3]
  return(tmp)
}
```


Aggregate field into grid cells, and remove this with fewer than 30 samples. We'll merge on seeding data.

```{r}
GridCells <- aggregate.field(seed.2018.dat,response='ControlRate')
GridCells <- GridCells[GridCells$Samples>30,]
#rename rate for simplicity
names(GridCells)[3] <- 'R18'
#and provide a relative rank
GridCells$RR <- GridCells$R18/max(GridCells$R18)

#Merge on row names, which will map to grid cell identifiers (row and column)
grid.2018.dat <- aggregate.field(harvest.2018.dat,response='Yield')
grid.2018.dat <- grid.2018.dat[grid.2018.dat$Samples>30,]
GridCells$Y18 <- grid.2018.dat[row.names(GridCells),'Yield']

#Since I create row names from row and column, we should be able to map yields correctly
cells.2017.dat <- aggregate.field(harvest.2017.dat,response='Yield')
GridCells$Y17 <- cells.2017.dat[row.names(GridCells),'Yield']

GridCells$Q18 <- floor(rank(GridCells$Y18)/max(rank(GridCells$Y18))/5)
GridCells$Q18[GridCells$Q18>5] = 5
GridCells$Q18 <- factor(GridCells$Q18*20)

tmp <- aggregate.field(seed.2020.dat,response='ControlRate')
GridCells$R20 <- tmp[row.names(GridCells),'ControlRate']
```

Add relative ranks and plot.

```{r,fig.width=8,fig.height=4.5}
GridCells$Y18r <- rank(GridCells$Y18)/max(rank(GridCells$Y18))
GridCells$Y17r <- rank(GridCells$Y17)/max(rank(GridCells$Y17))
GridCells$R18r <- rank(GridCells$R18)/max(rank(GridCells$R18))

GridMaps <- data.frame(Row=c(GridCells$Row,GridCells$Row,GridCells$Row),
                   Column=c(GridCells$Column, GridCells$Column, GridCells$Column),
                   Value=c(GridCells$Y18r,GridCells$Y17r,GridCells$R18r),
                   Map=c(rep('Y18',length(GridCells$Y18r)),
                         rep('Y17',length(GridCells$Y17r)),
                         rep('R18',length(GridCells$R18r))))
ggplot(GridMaps, aes(Column,Row)) + 
geom_point(aes(colour = Value),size=3) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Grid Data") + facet_wrap(~ Map)
```




```{r,eval=FALSE}
SeedMatrix <- matrix(0,nrow=max(GridCells$Row),ncol=max(GridCells$Column))
YieldMatrix <- matrix(0,nrow=max(GridCells$Row),ncol=max(GridCells$Column))
for(i in 1:max(GridCells$Row)) {
  for(j in 1:max(GridCells$Column)) {
      SeedMatrix[i,j] <- GridCells$R18[GridCells$Row==i & GridCells$Column==j]
      YieldMatrix[i,j] <- GridCells$Y18[GridCells$Row==i & GridCells$Column==j]
  }
}

col<- colorRampPalette(c(cbPalette[5],cbPalette[1],cbPalette[6]))(dim(GridCells)[1])
par(mfrow=c(1,2))
heatmap(t(SeedMatrix),Rowv=NA,Colv=NA,col=col)
heatmap(t(YieldMatrix),Rowv=NA,Colv=NA,col=col)
```


```{r,fig.width=8,fig.height=4.5}
ggplot(GridCells, aes(R18, Y18)) +
  geom_boxplot(aes(group = cut_width(R18, 500)), outlier.alpha = 0.1) +
geom_jitter(width = 100,alpha=0.5)
```


Model 0

$$
Y18_i = R18_i+e_i
$$

# Model 1

```{r}
Model1.lm <- lm(Y18 ~ 0+ R18 + I(R18^2) + I(R18^3),data=GridCells)
summary(Model1.lm)
anova(Model1.lm)
```


# Model 2

Consider adding yield from 2017 as a covariate - we are planting into the residuals from 2017.


```{r}
Model2.lm <- lm(Y18 ~ 0+Y17 + R18 + I(R18^2) + I(R18^3),data=GridCells)
summary(Model2.lm)
Model2.alt.lm <- lm(Y18 ~ 0+R18 + I(R18^2) + I(R18^3) + Y17,data=GridCells)
summary(Model2.alt.lm)


Model1.lm <- lm(Y18 ~ R18 + Y17, data=GridCells)
Model2.lm <- lm(Y18 ~ R18 + R18:Y17, data=GridCells)
Model3.lm <- lm(Y18 ~ R18 + Y17 + R18*Y17, data=GridCells)

summary(Model2.lm)
Model2.alt.lm <- lm(Y18 ~ Y17 + R18,data=GridCells)
summary(Model2.alt.lm)
Model2.2.lm <- lm(Y18 ~ Y17,data=GridCells)
summary(Model2.2.lm)
anova(Model2.2.lm)


Model3.alt.lm <- lm(Y18 ~ Y17 + R18 + R18*Y17,data=GridCells)
anova(Model2.lm)
anova(Model2.alt.lm)
anova(Model3.lm)
anova(Model3.alt.lm)
summary(Model3.lm)
summary(Model3.alt.lm)

AIC(Model1.lm,Model2.lm,Model3.lm)
BIC(Model1.lm,Model2.lm,Model3.lm)
```

```{r}
anova(Model1.lm,Model2.lm)
AIC(Model1.lm)
```

Note that there is not a unique decomposition of the SS for this model. This suggests seeding rate and previous seasons yield are confounded (or correlated). We need to consider a possible causative relationship.


```{r,fig.width=8,fig.height=4.5,eval=FALSE}

#ggpairs(GridCells[,c('Y18','R18','Y17','Q18')], aes(color=Q18,alpha = 0.4))
ggpairs(GridCells[,c('Y18','R18','Y17','Q18')], mapping = aes(color = Q18)) +
  scale_fill_manual(values = cbPalette)
```

```{r,fig.width=8,fig.height=4.5}
ggpairs(GridCells[,c('Y18','R18','Y17')]) +
  scale_fill_manual(values = cbPalette)
```


```{r}
Model.rate.lm <- lm(R18 ~ Y17,data=GridCells)
summary(Model.rate.lm)
anova(Model.rate.lm)
```

In fact, there is a greater degree of correlation between `R18` and `Y17`.




# DAG

```{r}
model1.dat <- GridCells[,c('Y18','R18','Y17')]
boot.strength(GridCells[,c('Y18','R18','Y17')],algorithm='pc.stable')
boot.strength(GridCells[,c('Y18','R18','Y17')], algorithm = "hc")
boot.strength(GridCells[,c('Y18','R18','Y17')], algorithm = "gs")

model1.dag <- model2network("[Y17][R18][Y18|R18:Y17]")
model2.dag <- model2network("[Y17][R18|Y17][Y18|R18]")
model3.dag <- model2network("[Y17][R18|Y17][Y18|R18:Y17]")
graphviz.plot(model1.dag)
graphviz.plot(model2.dag)
graphviz.plot(model3.dag)

#model1.dag
fit1 = bn.fit(model1.dag, model1.dat)

strength1 <- arc.strength(model1.dag, GridCells[,c('Y18','R18','Y17')])
strength1
strength.plot(model1.dag, strength1)
#model2.dag

fit2 = bn.fit(model2.dag, model1.dat)
fit2

strength2 <- arc.strength(model2.dag, model1.dat)
strength2
strength.plot(model2.dag, strength2)

fit3 = bn.fit(model3.dag, model1.dat)
fit3

strength3 <- arc.strength(model3.dag, model1.dat)

arc.strength(model3.dag, model1.dat)
arc.strength(model3.dag, model1.dat, criterion = 'aic-g')
arc.strength(model3.dag, model1.dat, criterion = 'bic-g')
arc.strength(model3.dag, model1.dat, criterion = 'loglik-g')


strength3
strength.plot(model3.dag, strength3)

bn.cv(model1.dat, model1.dag, algorithm.args = list(score = "aic"))
bn.cv(model1.dat, model2.dag, algorithm.args = list(score = "aic"))
bn.cv(model1.dat, model3.dag, algorithm.args = list(score = "aic"))

bf.strength1 <- bf.strength(model1.dag, model1.dat)
bf.strength2 <- bf.strength(model2.dag, model1.dat)
bf.strength3 <- bf.strength(model3.dag, model1.dat)

strength.plot(model1.dag, bf.strength1)
strength.plot(model2.dag, bf.strength2)
strength.plot(model3.dag, bf.strength3)

averaged.network(bf.strength1)
averaged.network(bf.strength2)
averaged.network(bf.strength3)

bf.strength1
bf.strength2 
bf.strength3

score(model1.dag,model1.dat,type = 'aic-g')
score(model2.dag,model1.dat,type = 'aic-g')
score(model3.dag,model1.dat,type = 'aic-g')

score(model1.dag,model1.dat,type = 'bic-g')
score(model2.dag,model1.dat,type = 'bic-g')
score(model3.dag,model1.dat,type = 'bic-g')

score(model1.dag,model1.dat,type = 'loglik-g')
score(model2.dag,model1.dat,type = 'loglik-g')
score(model3.dag,model1.dat,type = 'loglik-g')

score(model1.dag,model1.dat,type = 'bge')
score(model2.dag,model1.dat,type = 'bge')
score(model3.dag,model1.dat,type = 'bge')
```


```{r}
learned.dag = iamb(GridCells[,c('Y18','R18','Y17')])
graphviz.plot(learned.dag)
graphviz.plot(pc.stable(GridCells[,c('Y18','R18','Y17')]))
#learned.dag = set.arc(learned.dag, from = "Y17", to = "Y18")
#learned.dag = set.arc(learned.dag, from = "R18", to = "Y18")
#graphviz.plot(learned.dag)
#fit1 = bn.fit(learned.dag, GridCells[,c('Y18','R18','Y17')])
#fit1

#learned.dag = set.arc(learned.dag, from = "Y17", to = "R18")
#graphviz.plot(learned.dag)
#fit1 = bn.fit(learned.dag, GridCells[,c('Y18','R18','Y17')])
#fit1
```


# Model 4

```{r}
tmp <- aggregate.field(seed.2018.dat,response='AppliedRate')
GridCells$AR18 <- tmp[row.names(GridCells),'AppliedRate']
```


```{r,fig.width=8,fig.height=4.5,eval=FALSE}
ggpairs(GridCells[,c('Y18','R18','Y17','AR18','Q18')], mapping = aes(color = Q18)) +
  scale_fill_manual(values = cbPalette)
```

```{r,fig.width=8,fig.height=4.5}
ggpairs(GridCells[,c('Y18','R18','Y17','AR18')]) +
  scale_fill_manual(values = cbPalette)
```

```{r}
model4.dag <- model2network("[Y17][R18|Y17][AR18|R18][Y18|AR18:Y17]")
graphviz.plot(model4.dag)
fit1 = bn.fit(model4.dag, GridCells[,c('Y18','R18','Y17','AR18')])
fit1
strength4 <- arc.strength(model4.dag, GridCells[,c('Y18','R18','Y17','AR18')])

strength.plot(model4.dag, strength4)


bf.strength4 <- bf.strength(model4.dag, GridCells[,c('Y18','R18','Y17','AR18')])
strength.plot(model4.dag, bf.strength4)
averaged.network(bf.strength4)
plot(bf.strength4)

strength4
bf.strength4
```

```{r}
Model4.lm <- lm(Y18 ~ R18 + Y17 + AR18,data=GridCells)
summary(Model4.lm)
anova(Model4.lm)

Model4.1.lm <- lm(AR18 ~ R18,data=GridCells)
summary(Model4.1.lm)

Model4.2.lm <- lm(R18 ~ Y17,data=GridCells)
summary(Model4.2.lm)

Model4.3.lm <- lm(Y17 ~ 1,data=GridCells)
summary(Model4.3.lm)

Model4.4.lm <- lm(Y18 ~ AR18 + Y17,data=GridCells)
summary(Model4.4.lm)
```


```{r}
learned.dag = iamb(GridCells[,c('Y18','R18','Y17','AR18')])
graphviz.plot(learned.dag)
#learned.dag = set.arc(learned.dag, from = "Y17", to = "Y18")
##learned.dag = set.arc(learned.dag, from = "Y17", to = "R18")
#learned.dag = set.arc(learned.dag, from = "R18", to = "Y18")
#graphviz.plot(learned.dag)
#fit1 = bn.fit(learned.dag, GridCells[,c('Y18','R18','Y17')])
#fit1
```


## Model 5 and 6

We want to incorporate information about prior years' yields.

```{r,fig.width=8,fig.height=4.5}
harvest.2016.dat$YieldRank <- rank(harvest.2016.dat$Yield)
harvest.2016.dat$YieldRank <- harvest.2016.dat$YieldRank/max(harvest.2016.dat$YieldRank)
harvest.2015.dat$YieldRank <- rank(harvest.2015.dat$Yield)
harvest.2015.dat$YieldRank <- harvest.2015.dat$YieldRank/max(harvest.2015.dat$YieldRank)
harvest.2013.dat$YieldRank <- rank(harvest.2013.dat$Yield)
harvest.2013.dat$YieldRank <- harvest.2013.dat$YieldRank/max(harvest.2013.dat$YieldRank)

Maps5 <- data.frame(Longitude=c(harvest.2018.dat$Longitude,
                                harvest.2017.dat$Longitude,
                                harvest.2016.dat$Longitude,
                                harvest.2015.dat$Longitude,
                                seed.2018.dat$Longitude),
                   Latitude=c(harvest.2018.dat$Latitude,
                              harvest.2017.dat$Latitude,
                              harvest.2016.dat$Latitude,
                              harvest.2015.dat$Latitude,
                              seed.2018.dat$Latitude),
                   Value=c(harvest.2018.dat$YieldRank,
                           harvest.2017.dat$YieldRank,
                           harvest.2016.dat$YieldRank,
                           harvest.2015.dat$YieldRank,
                           seed.2018.dat$RateRank),
                   Map=c(rep('Y18',length(harvest.2018.dat$YieldRank)),
                         rep('Y17',length(harvest.2017.dat$YieldRank)),
                         rep('Y16',length(harvest.2016.dat$YieldRank)),
                         rep('Y15',length(harvest.2015.dat$YieldRank)),
                         rep('R18',length(seed.2018.dat$RateRank))))
ggplot(Maps5, aes(Longitude,Latitude)) + 
geom_point(aes(colour = Value),size=.5) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Multiple Years") + facet_wrap(~ Map)
```




```{r,fig.width=8,fig.height=4.5,eval=FALSE}


model6_dag <- dagify(Y18 ~ AR18,
       Y18 ~ Y17,
       AR18 ~ R18,
       R18 ~ RS,
       Y18 ~ RS,
       Y17 ~ Y16,
       Y16 ~ Y15,
       Y15 ~ RS,
       Y16 ~ RS,
       Y17 ~ RS,
       labels = c("Y18" = "Yield\n 2018", 
                  "R18" = "Control Rate",
                  "Y17" = "Yield \n2017"),
       latent = "RS",
       #exposure = "smoking",
       outcome = "Y18")
#ggdag(model6_dag, text = TRUE, use_labels = "label")
ggdag(model6_dag, text = TRUE)
```


```{r,fig.width=7,fig.height=5,eval=FALSE}
harvest.2019.dat$YieldRank <- rank(harvest.2019.dat$Yield)
harvest.2019.dat$YieldRank <- harvest.2019.dat$YieldRank/max(harvest.2019.dat$YieldRank)
ggplot(harvest.2019.dat, aes(Longitude,Latitude)) + 
geom_point(aes(colour = YieldRank),size=1) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue) +
labs(colour = "Yield (Relative Rank)", x="Easting", y="Northing", title = "2019")
```

```{r}
cells.2013.dat <- aggregate.field(harvest.2013.dat,response='Yield')
cells.2015.dat <- aggregate.field(harvest.2015.dat,response='Yield')
cells.2016.dat <- aggregate.field(harvest.2016.dat,response='Yield')

cells.2019.dat <- aggregate.field(harvest.2019.dat,response='Yield')
cells.2020.dat <- aggregate.field(harvest.2020.dat,response='Yield')

GridCells$Y13 <- cells.2013.dat[row.names(GridCells),'Yield']
GridCells$Y15 <- cells.2015.dat[row.names(GridCells),'Yield']
GridCells$Y16 <- cells.2016.dat[row.names(GridCells),'Yield']
GridCells$Y19 <- cells.2019.dat[row.names(GridCells),'Yield']
GridCells$Y20 <- cells.2019.dat[row.names(GridCells),'Yield']
```


```{r,fig.width=8,fig.height=4}
ggpairs(GridCells[,c('Y18','AR18','R18','Y17','Y16','Y15')], aes(alpha = 0.4))
```



```{r}
Model5.1.lm <- lm(AR18 ~ R18,data=GridCells)
summary(Model5.1.lm)

Model5.2.lm <- lm(R18 ~ Y15 + Y16 + Y17,data=GridCells)
summary(Model5.2.lm)

Model5.3.lm <- lm(Y15 ~ 1,data=GridCells)
summary(Model5.3.lm)

Model5.4.lm <- lm(Y16 ~ Y15,data=GridCells)
summary(Model5.4.lm)

Model5.5.lm <- lm(Y17 ~ Y16,data=GridCells)
summary(Model5.5.lm)

Model5.6.lm <- lm(Y18 ~ AR18 + Y17,data=GridCells)
summary(Model5.6.lm)
```

```{r}
model5.dat <- GridCells[,c('Y18',"R18",'AR18',"Y17",'Y16','Y15')]
model5.dag <- model2network("[Y15][Y16|Y15][Y17|Y16][R18|Y17:Y16:Y15][AR18|R18][Y18|AR18:Y17]")
#graphviz.plot(model5.dag,layout = "dot")
#graphviz.plot(model5.dag,layout = "fdp")
graphviz.plot(model5.dag,layout = "circo")
model5.fit = bn.fit(model5.dag, model5.dat)
model5.fit
#bn.fit.qqplot(model5.fit)
strength5 <- arc.strength(model5.dag, model5.dat)

strength.plot(model5.dag, strength5,layout = "circo")

bf.strength5 <- bf.strength(model5.dag, model5.dat)

strength.plot(model5.dag, bf.strength5)
averaged.network(bf.strength5)
plot(bf.strength5)

strength5
bf.strength5

arc.strength(model5.dag, model5.dat, criterion = 'aic-g')

```

```{r}
model6.dat <- GridCells[,c('Y18',"R18",'AR18',"Y17",'Y16','Y15')]
model6.dag <- model2network("[Y15][Y16|Y15][Y17|Y16:Y15][R18|Y17:Y16:Y15][AR18|R18][Y18|AR18:Y17:Y16:Y15]")
model6.dag
model6.fit = bn.fit(model6.dag, model6.dat)
model6.fit
#bn.fit.qqplot(model6.fit)
strength6 <- arc.strength(model6.dag, model6.dat)

strength.plot(model6.dag, strength6)

bf.strength6 <- bf.strength(model6.dag, model6.dat)

strength.plot(model6.dag, bf.strength6)
averaged.network(bf.strength6)
plot(bf.strength6)

strength6
bf.strength6

arc.strength(model6.dag, model6.dat, criterion = 'aic-g')

bn.cv(model5.dat, model5.dag)
bn.cv(model6.dat, model6.dag)


score(model5.dag,model5.dat,type = 'bge')
score(model5.dag,model5.dat,type = 'aic-g')
score(model6.dag,model6.dat,type = 'bge')
score(model6.dag,model6.dat,type = 'aic-g')

BF(model5.dag, model6.dag, model6.dat)

```

```{r,eval=TRUE}
#VarCorr(default.gam)
#learned.dag = iamb(GridCells[,c('Y18',"R18","Y17","Y16","Y15")])
learned.dag = iamb(GridCells[,c('Y18','AR18',"R18","Y17","Y16","Y15")])
graphviz.plot(learned.dag)
```

```{r,eval=FALSE}
# learned.dag = set.arc(learned.dag, from = "A", to = "O") 
#learned.dag = drop.arc(learned.dag, from = "E", to = "O") 
#learned.dag = reverse.arc(learned.dag, from = "R", to = "E")

undirected.arcs(learned.dag)

learned.fit = bn.fit(learned.dag, GridCells[,c('Y18',"R18","Y17","Y16","Y15","Y13")])
learned.fit

learned.dag = drop.arc(learned.dag, from = "Y18", to = "Y13")
learned.dag = drop.arc(learned.dag, from = "Y15", to = "Y13")
learned.dag = drop.arc(learned.dag, from = "Y16", to = "Y13")
learned.dag = set.arc(learned.dag, from = "Y13", to = "Y17") 
graphviz.plot(learned.dag)
learned.fit = bn.fit(learned.dag, GridCells[,c('Y18',"R18","Y17","Y16","Y15","Y13")])
learned.fit

bn.fit.qqplot(learned.fit)
bn.fit.histogram(learned.fit)
bn.fit.xyplot(learned.fit)
## for discrete (multinomial and ordinal) Bayesian networks.
#bn.fit.barchart(learned.fit)
#bn.fit.dotplot(learned.fit)
```


```{r,eval=TRUE}
#VarCorr(default.gam)
#learned.dag = iamb(GridCells[,c('Y18',"R18","Y17","Y16","Y15")])
learned.dag = iamb(GridCells[,c('R20','Y19','Y18','AR18',"R18","Y17","Y16","Y15","Y13")])
graphviz.plot(learned.dag)
```

# Repeat with ranks on yields

```{r}
GridRanks <- GridCells
GridRanks$Y20 <- rank(GridCells$Y20)/max(rank(GridCells$Y20))
GridRanks$Y19 <- rank(GridCells$Y19)/max(rank(GridCells$Y19))
GridRanks$Y18 <- rank(GridCells$Y18)/max(rank(GridCells$Y18))
GridRanks$Y17 <- rank(GridCells$Y17)/max(rank(GridCells$Y17))
GridRanks$Y16 <- rank(GridCells$Y16)/max(rank(GridCells$Y16))
GridRanks$Y15 <- rank(GridCells$Y15)/max(rank(GridCells$Y15))
GridRanks$Y13 <- rank(GridCells$Y13)/max(rank(GridCells$Y13))
GridRanks$AR18 <- rank(GridCells$AR18)/max(rank(GridCells$AR18))
GridCells$RS18 <- apply(GridRanks[,c('Y17','Y16','Y15','Y13')],1,mean)
```


```{r}
model5.dat <- GridRanks[,c('Y18',"R18",'AR18',"Y17",'Y16','Y15')]
model5.dag <- model2network("[Y15][Y16|Y15][Y17|Y16][R18|Y17:Y16:Y15][AR18|R18][Y18|AR18:Y17]")
#graphviz.plot(model5.dag,layout = "dot")
#graphviz.plot(model5.dag,layout = "fdp")
graphviz.plot(model5.dag,layout = "circo")
model5.fit = bn.fit(model5.dag, model5.dat)
model5.fit
#bn.fit.qqplot(model5.fit)
strength5 <- arc.strength(model5.dag, model5.dat)
strength5
strength.plot(model5.dag, strength5,layout = "circo")
```


```{r}
model6.dat <- GridRanks[,c('Y18',"R18",'AR18',"Y17",'Y16','Y15')]
model6.dag <- model2network("[Y15][Y16|Y15][Y17|Y16:Y15][R18|Y17:Y16:Y15][AR18|R18][Y18|AR18:Y17:Y16:Y15]")
model6.dag
model6.fit = bn.fit(model6.dag, model6.dat)
model6.fit
#bn.fit.qqplot(model6.fit)
strength5 <- arc.strength(model6.dag, model6.dat)
strength5
strength.plot(model6.dag, strength5)

bn.cv(model5.dat, model5.dag)
bn.cv(model6.dat, model6.dag)
```

# Model 7 and  8

Model a response surface
```{r}
model7.dat <- GridCells[,c('Y18',"R18",'AR18',"Y17",'Y16','Y15','Y13','RS18')]
model7.dag <- model2network("[Y13][Y15][RS18|Y13:Y15:Y16:Y17][Y16|Y15][Y17|Y16][R18|Y17:RS18][AR18|R18][Y18|AR18:Y17:RS18]")
model7.dag
model7.fit = bn.fit(model7.dag, model7.dat)
model7.fit
#bn.fit.qqplot(model6.fit)
strength7 <- arc.strength(model7.dag, model7.dat)
strength7
strength.plot(model7.dag, strength7)

model8.dat <- GridCells[,c('Y18',"R18",'AR18',"Y17",'Y16','Y15','Y13')]
model8.dag <- model2network("[Y15][Y13][Y16|Y15][Y17|Y16:Y15][R18|Y17:Y16:Y15:Y13][AR18|R18][Y18|AR18:Y17:Y16:Y15:Y13]")
model8.dag
model8.fit = bn.fit(model8.dag, model8.dat)
model8.fit
#bn.fit.qqplot(model6.fit)
strength8 <- arc.strength(model8.dag, model8.dat)
strength8
strength.plot(model8.dag, strength8)

bn.cv(model7.dat, model7.dag)
bn.cv(model8.dat, model8.dag)
```


```{r}
model5.rank.fit = bn.fit(model5.dag, GridRanks[,c('Y18','AR18',"R18","Y17",'Y16','Y15')])
model5.rank.fit
bn.fit.qqplot(model5.rank.fit)
```

```{r}
learned.rank.dag = iamb(GridRanks[,c('Y18',"R18","Y17","Y16","Y15","Y13")])
undirected.arcs(learned.rank.dag)
graphviz.plot(learned.rank.dag)
#learned.rank.fit = bn.fit(learned.rank.dag, mat)
#learned.rank.fit
```






```{r,fig.width=8,fig.height=4}
ggpairs(GridCells[,c('Y19','Y18','AR18','Y17','Y16','Y15','Y13')], aes(alpha = 0.4))
```


```{r,fig.width=8,fig.height=4}
ggpairs(GridCells[,c('R20','Y19','R18','Y18','Y17','Y16','Y15','Y13')], aes(alpha = 0.4))
```


```{r}
#model9.dat <- GridCells[,c('R20','Y19','Y18',"R18",'AR18',"Y17",'Y16','Y15')]
#model9.dag <- model2network("[Y15][Y16|Y15][Y17|Y16:Y15][R18|Y17:Y16:Y15][AR18|R18][Y18|AR18:Y17:Y16:Y15][Y19|Y18:Y17:Y16:Y15][R20|Y19:Y18:Y17:Y16:Y15]")

model9.dat <- GridCells[,c('Y20','R20','Y19','Y18',"R18",'AR18',"Y17",'Y16','Y15')]
model9.dag <- model2network("[Y15][Y16|Y15][Y17|Y16:Y15][R18|Y17:Y16:Y15][AR18|R18][Y18|AR18:Y17:Y16:Y15][Y19|Y18:Y17:Y16:Y15][R20|Y19:Y18:Y17:Y16:Y15][Y20|R20:Y19:Y18:Y17:Y16:Y15]")


model9.fit = bn.fit(model9.dag, model9.dat)
model9.fit
bn.fit.qqplot(model9.fit)
strength9 <- arc.strength(model9.dag, model9.dat)
strength9
strength.plot(model9.dag, strength9)

bf.strength9 <- bf.strength(model9.dag, model9.dat)
bf.strength9
strength.plot(model9.dag, bf.strength9)
averaged.network(bf.strength9)
plot(bf.strength9)

```

```{r,fig.width=8,fig.height=4.5}
harvest.2016.dat$YieldRank <- rank(harvest.2016.dat$Yield)
harvest.2016.dat$YieldRank <- harvest.2016.dat$YieldRank/max(harvest.2016.dat$YieldRank)
harvest.2015.dat$YieldRank <- rank(harvest.2015.dat$Yield)
harvest.2015.dat$YieldRank <- harvest.2015.dat$YieldRank/max(harvest.2015.dat$YieldRank)
harvest.2013.dat$YieldRank <- rank(harvest.2013.dat$Yield)
harvest.2013.dat$YieldRank <- harvest.2013.dat$YieldRank/max(harvest.2013.dat$YieldRank)

harvest.2019.dat$YieldRank <- rank(harvest.2019.dat$Yield)
harvest.2019.dat$YieldRank <- harvest.2019.dat$YieldRank/max(harvest.2019.dat$YieldRank)

Maps4 <- data.frame(Longitude=c(#harvest.2018.dat$Longitude,
                                harvest.2017.dat$Longitude,
                                seed.2018.dat$Longitude,
                                harvest.2019.dat$Longitude,
                                seed.2020.dat$Longitude),
                   Latitude=c(#harvest.2018.dat$Latitude,
                              harvest.2017.dat$Latitude,
                              seed.2018.dat$Latitude,
                              harvest.2019.dat$Latitude,
                              seed.2020.dat$Latitude),
                   Value=c(#harvest.2018.dat$YieldRank,
                           harvest.2017.dat$YieldRank,
                           seed.2018.dat$RateRank,
                           harvest.2019.dat$YieldRank,
                           seed.2020.dat$RateRank),
                   Map=c(#rep('Harvest 2018',length(harvest.2018.dat$YieldRank)),
                         rep("Harvest 2017",length(harvest.2017.dat$YieldRank)),
                         rep("Seed 2018",length(seed.2018.dat$RateRank)),
                         rep('Harvest 2019',length(harvest.2019.dat$YieldRank)),
                         rep('Seed 2020',length(seed.2020.dat$RateRank))))
ggplot(Maps4, aes(Longitude,Latitude)) + 
geom_point(aes(colour = Value),size=1) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Two Seeding Maps") + facet_wrap(~ Map)
```

# Proposed

```{r,fig.width=8,fig.height=4.5}
seed.2020.dat$Pass = floor(seed.2020.dat$Longitude/25)
CR <- sample(unique(seed.2020.dat$Pass),10) 
CR = 20 + 2*(1:5)
seed.2020.dat$CR <- seed.2020.dat$Pass %in% CR
seed.2020.dat$Proposed <- seed.2020.dat$ControlRate
seed.2020.dat$Proposed[seed.2020.dat$CR] <- median(seed.2020.dat$ControlRate)

seed.2020.dat$PS <- rank(seed.2020.dat$Proposed)
seed.2020.dat$PS <- seed.2020.dat$PS/max(seed.2020.dat$PS)

ggplot(seed.2020.dat, aes(Longitude,Latitude)) + 
geom_point(aes(colour = PS),size=1) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Two Seeding Maps")

Maps2 <- data.frame(Longitude=c(seed.2020.dat$Longitude,
                                seed.2020.dat$Longitude),
                   Latitude=c(seed.2020.dat$Latitude,
                              seed.2020.dat$Latitude),
                   Value=c(seed.2020.dat$RateRank,
                           seed.2020.dat$PS),
                   Map=c(rep('Actual 2020',length(seed.2020.dat$RateRank)),
                         rep('Proposed 2020',length(seed.2020.dat$PS))))
ggplot(Maps2, aes(Longitude,Latitude)) + 
geom_point(aes(colour = Value),size=1) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Experimental Design") + facet_wrap(~ Map)

```


```{r,fig.width=8,fig.height=4.5}
harvest.2016.dat$YieldRank <- rank(harvest.2016.dat$Yield)
harvest.2016.dat$YieldRank <- harvest.2016.dat$YieldRank/max(harvest.2016.dat$YieldRank)
harvest.2015.dat$YieldRank <- rank(harvest.2015.dat$Yield)
harvest.2015.dat$YieldRank <- harvest.2015.dat$YieldRank/max(harvest.2015.dat$YieldRank)
harvest.2013.dat$YieldRank <- rank(harvest.2013.dat$Yield)
harvest.2013.dat$YieldRank <- harvest.2013.dat$YieldRank/max(harvest.2013.dat$YieldRank)

harvest.2019.dat$YieldRank <- rank(harvest.2019.dat$Yield)
harvest.2019.dat$YieldRank <- harvest.2019.dat$YieldRank/max(harvest.2019.dat$YieldRank)

Maps4 <- data.frame(Longitude=c(#harvest.2018.dat$Longitude,
                                harvest.2017.dat$Longitude,
                                seed.2018.dat$Longitude,
                                harvest.2019.dat$Longitude,
                                seed.2020.dat$Longitude),
                   Latitude=c(#harvest.2018.dat$Latitude,
                              harvest.2017.dat$Latitude,
                              seed.2018.dat$Latitude,
                              harvest.2019.dat$Latitude,
                              seed.2020.dat$Latitude),
                   Value=c(#harvest.2018.dat$YieldRank,
                           harvest.2017.dat$YieldRank,
                           seed.2018.dat$RateRank,
                           harvest.2019.dat$YieldRank,
                           seed.2020.dat$RateRank),
                   Map=c(#rep('Harvest 2018',length(harvest.2018.dat$YieldRank)),
                         rep("Harvest 2017",length(harvest.2017.dat$YieldRank)),
                         rep("Seed 2018",length(seed.2018.dat$RateRank)),
                         rep('Harvest 2019',length(harvest.2019.dat$YieldRank)),
                         rep('Seed 2020',length(seed.2020.dat$RateRank))))
ggplot(Maps4, aes(Longitude,Latitude)) + 
geom_point(aes(colour = Value),size=1) + 
scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = 0.5) +
labs(colour = "Relative Rank", x="Easting", y="Northing", title = "Two Seeding Maps") + facet_wrap(~ Map)
```

```{r,fig.width=8,fig.height=4.5}
ggplot(GridCells, aes(Y18,Y19)) + geom_point(aes(color=R18), size=3) + scale_colour_gradient2(low=vermillion, mid=yellow, high=blue, midpoint = mean(GridCells$R18)) +
 geom_smooth(se=FALSE,method = 'lm',color=grey)
```

# Joining maps

See CausalInferenceAFullMaps


```{r,eval=TRUE}
#VarCorr(default.gam)
#learned.dag = iamb(GridCells[,c('Y18',"R18","Y17","Y16","Y15")])
learned.dag = iamb(GridCells[,c('Y18','AR18',"R18","Y17","Y16","Y15","Y13")])
graphviz.plot(learned.dag)
```

```{r,eval=FALSE}
model9.dag <- model2network("[Y15][Y16|Y15][Y17|Y16:Y15][R18|Y17:Y16:Y15][AR18|R18][Y18|AR18:Y17:Y16:Y15][Y19|Y18][R20|Y19:Y18:Y17:Y16:Y15]")

model9.fit = bn.fit(model9.dag, mapped.dat[,c('R20','Y19','Y18',"R18",'AR18',"Y17",'Y16','Y15')])
model9.fit
bn.fit.qqplot(model9.fit)
strength9 <- arc.strength(model9.dag, mapped.dat[,c('R20','Y19','Y18',"R18",'AR18',"Y17",'Y16','Y15')])
strength9
strength.plot(model9.dag, strength9)
```

```{r}
model9.dat <- GridCells[,c('Y20','R20','Y19','Y18',"R18",'AR18',"Y17",'Y16','Y15')]
model9.dag <- model2network("[Y15][Y16|Y15][Y17|Y16:Y15][R18|Y17:Y16:Y15][AR18|R18][Y18|AR18:Y17:Y16:Y15][Y19|Y18:Y17:Y16:Y15][R20|Y19:Y18:Y17:Y16:Y15][Y20|R20:Y19:Y18:Y17:Y16:Y15]")

model9.fit = bn.fit(model9.dag, model9.dat)
model9.fit
bn.fit.qqplot(model9.fit)

bf.strength9 <- bf.strength(model9.dag, model9.dat)

strength.plot(model9.dag, bf.strength9)
averaged.network(bf.strength9)
plot(bf.strength9)

bf.strength9 
```


# 2019 Response

```{r,fig.width=8,fig.height=4.5}
GridLong <- reshape(GridCells, 
  varying = c('Y18','Y17','Y16','Y15','Y13','Y19'), 
  v.names = "Yield",
  timevar = "Year", 
  times = c(2018,2017,2016,2015,2013,2019), 
  direction = "long")
plot(Yield~Column,data=GridLong)
GridLong$Year <- factor(GridLong$Year)
ggplot(GridLong, aes(Column,Yield,color=Year)) + scale_colour_manual(values=cbPalette) +
geom_point(size=2) + geom_smooth(se = FALSE)
```

```{r,fig.width=8,fig.height=4.5}
GridRankLong <- reshape(GridRanks, 
  varying = c('Y18','Y17','Y16','Y15','Y13','Y19'), 
  v.names = "Yield",
  timevar = "Year", 
  times = c(2018,2017,2016,2015,2013,2019), 
  direction = "long")
plot(Yield~Column,data=GridLong)
GridRankLong$Year <- factor(GridRankLong$Year)
ggplot(GridRankLong, aes(Column,Yield,color=Year)) + scale_colour_manual(values=cbPalette) +
geom_point(size=2) + geom_smooth(se = FALSE)
```

```{r,fig.width=8,fig.height=4.5}
GridRankLong$Pass <- factor(floor(GridRankLong$Column/4))
ggplot(GridRankLong, aes(Pass, Yield,color=Year)) +
  geom_boxplot(aes(color = Year), outlier.alpha = 0.4) + geom_smooth(se = FALSE)
#geom_jitter(width = 2,alpha=0.4,size=2) + scale_colour_manual(values=cbPalette)
```


# Response plots

Plot achieved yield against actual applied rates versus prescribed rates.



TODO : optimize grid size

# Generalized Additive Model

```{r}

red.gam <- gam(Y18 ~  Y17, data=GridCells)
lm.gam <- gam(Y18 ~ R18 + Y17, data=GridCells)
gam.check(red.gam)
gam.check(lm.gam)
```

```{r}
red.s.gam <- gam(Y18 ~ s(Y17), data=GridCells)
s.gam <- gam(Y18 ~ s(R18) + s(Y17), data=GridCells)
gam.check(red.s.gam)
gam.check(s.gam)
```
```{r}
k9.gam <- gam(Y18 ~ s(R18,k=8) + s(Y17,k=8), data=GridCells)
```

```{r}
summary(lm.gam)
summary(s.gam)
summary(k9.gam)

anova(lm.gam)
anova(s.gam)
anova(k9.gam)
```

```{r}
red.gam <- gam(Y18 ~  Y17 + Y16 + Y15 + Y13, data=GridCells)
lm.gam <- gam(Y18 ~ R18 + Y17 + Y16 + Y15 + Y13, data=GridCells)
gam.check(red.gam)
gam.check(lm.gam)
```

```{r}
red.s.gam <- gam(Y18 ~ s(Y17) + s(Y16) + s(Y15) + s(Y13), data=GridCells)
s.gam <- gam(Y18 ~ s(R18) + s(Y17) + s(Y16) + s(Y15) + s(Y13), data=GridCells)
gam.check(red.s.gam)
gam.check(s.gam)
```
```{r}
k9.gam <- gam(Y18 ~ s(R18,k=9) + s(Y17,k=9) + s(Y16,k=9) + s(Y15,k=9) + s(Y13,k=9), data=GridCells)
```

```{r}
summary(lm.gam)
summary(s.gam)
summary(k9.gam)

anova(lm.gam)
anova(s.gam)
anova(k9.gam)
```

```{r}
plot(lm.gam,pages=1,residuals=TRUE,all.terms=TRUE,shade=TRUE,shade.col=2)
plot(s.gam,pages=1,residuals=TRUE,all.terms=TRUE,shade=TRUE,shade.col=2)
plot(k9.gam,pages=1,residuals=TRUE,all.terms=TRUE,shade=TRUE,shade.col=2)
```

```{r,eval=FALSE}
seed.2018.dat <- grid.field(seed.2018.dat)
cells.seed.dat$AR18 <- aggregate(AppliedRate ~ Row + Column,data=mapped.dat,FUN=mean,na.rm=TRUE)$AppliedRate
```







```{r, echo=FALSE,eval=FALSE}
library(DiagrammeR)
DiagrammeR::grViz("digraph rmarkdown {Rate -> Yield}", height=200)
```

```{r, echo=FALSE,eval=FALSE}
DiagrammeR::grViz("digraph rmarkdown {
                 Fertility -> Rate  
                 Fertility -> Yield}", height=200)
```

```{r, echo=FALSE,eval=FALSE}
DiagrammeR::grViz('digraph rmarkdown {
                 Fertility -> Rate  
                 Rate -> Yield [ style=dashed label=indirect ]
                 Fertility -> Yield}', height=200)
```


```{r, echo=FALSE,eval=FALSE}
DiagrammeR::grViz('digraph rmarkdown {
                 Fertility -> ControlRate  
                 ControlRate -> AppliedRate [style=dashed]
                 AppliedRate -> Yield [style=dashed]
                 Fertility -> Yield}', height=200)
```



```{r}
Model2_dag <- dagify(Y18 ~ R18,
       R18 ~ Y17,
       Y18 ~ Y17,
       labels = c("Y18" = "Yield\n 2018", 
                  "R18" = "Seeding Rate",
                  "Y17" = "Yield \n2017"),
       #latent = "unhealthy",
       #exposure = "smoking",
       outcome = "Y18")

ggdag(Model2_dag, text = FALSE, use_labels = "label")
```




```{r,fig.width=8,fig.height=4.5}
Model2_dag <- dagify(Y18 ~ R18,
       Y18 ~ Y17,
       labels = c("Y18" = "Yield\n 2018", 
                  "R18" = "Seeding Rate",
                  "Y17" = "Yield \n2017"),
       #latent = "unhealthy",
       #exposure = "smoking",
       outcome = "Y18")

ggdag(Model2_dag, text = FALSE, use_labels = "label")
```




